### 2025-09-25 paper/sec/introduction.tex
comment
rewrite of the RQs
Improve Fig 1 caption.
Tweak.

Making sense of and verifying data-driven claims is hard even with open and accessible data. A key challenge
lies in tracing specific claims back to the relevant data. In peer review, for example, empirical claims
typically lack author-supplied links to data, making them hard for reviewers to check
directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to simple
errors in data management or analysis~\citep{hu25}. The use of large language models (LLMs) to interpret
scholarly documents has seen considerable attention recently, from fact-checking~\citep{abu-ahmad25} to
interpretation of charts and figures~\citep{roberts24}, but current LLM interfaces do not support direct
interrogation of visual or other outputs for traceability to inputs.

@@

\begin{figure}%[h]
    \centering
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
    \vspace{1mm}
    \hrule
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
    \caption{\emph{Transparent document} example, showing data-driven natural language (top
    vs.~bottom)}
    \label{fig:scigen-example-website}
\end{figure}

@@

\subsection{Target idioms of natural language}

>>>>

When attempting to interpret or verify data-driven claims, a key challenge lies in tracing specific claims
back to the relevant data. In peer review, for example, empirical claims typically lack author-supplied links
to data, making them hard for reviewers to check directly~\citep{weber20}. Paper retractions, meanwhile, are
often attributable not to fraud, but to simple errors in data management or analysis~\citep{hu25}. The use of
large language models (LLMs) to interpret scholarly documents has seen considerable attention recently, from
fact-checking~\citep{abu-ahmad25} to interpretation of charts and figures~\citep{roberts24}, but current LLM
interfaces do not support direct interrogation of visual or other outputs for traceability to inputs.

@@

\begin{figure}%[h]
    \centering
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
    \vspace{1mm}
    \hrule
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
    \caption{Two versions of a transparent document, showing text fragments linked to data}
    \label{fig:scigen-example-website}
\end{figure}

@@

\subsection{Target idioms of natural language}
\cdnote{Can we make this a section? intro is getting too long}
